\documentclass[11pt]{article}

% ---- packages ----
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage[margin=1in]{geometry}
\usepackage{natbib}

% ---- helper for plan notes ----
\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
\newcommand{\plan}[1]{\textcolor{blue}{\textit{#1}}}

\title{Yedi: A Multi-Dimensional Strategic Reasoning Benchmark\\for Large Language and Vision-Language Models}

\author{
  \todo{Authors and affiliations}
}

\date{}

\begin{document}
\maketitle

\begin{center}
\small
Code and benchmark: \url{https://github.com/yilmazozgur/yedi-v4}\\
Video walkthroughs of the original mobile game: \url{https://www.youtube.com/@mindfreegames7459}
\end{center}

% ======================================================================
\begin{abstract}
\plan{%
  3--4 sentences. Key points to hit:
  \begin{itemize}
    \item Existing game benchmarks test single-axis reasoning (math, spatial, language).
      Real decisions require simultaneous optimization across multiple cognitive
      dimensions under resource constraints.
    \item We present Yedi, a configurable card-game benchmark where merge outcomes
      are determined by multiplicative interaction across 1--4 independent
      dimensions (numerical, visual, spatial, verbal). One bad dimension
      ruins the compound gain --- models must optimize the worst-case factor,
      not the average.
    \item We provide a Gymnasium environment with 40+ difficulty configurations,
      metadata and screenshot observation modes, and three baseline agents
      (random, greedy, LLM/VLM).
    \item Early results: frontier LLMs surpass a strong greedy baseline in
      metadata mode. Comparing metadata vs.\ vision modes on identical
      configs quantifies a ``vision tax'' that isolates perceptual grounding
      cost from strategic reasoning ability.
  \end{itemize}
}
\end{abstract}

% ======================================================================
\section{Introduction}
\label{sec:intro}

\plan{%
  Motivation paragraph (why another benchmark?):
  \begin{itemize}
    \item LLMs/VLMs are increasingly evaluated on game-playing tasks
      (Tetris, 2048, chess, Go, Atari). These are valuable but typically
      test a single axis --- spatial planning, arithmetic, pattern recognition.
    \item Real-world decisions (medical, financial, engineering) require
      balancing multiple incommensurable factors simultaneously. A treatment
      that is optimal for one organ system but harmful to another is not a
      good treatment. Current benchmarks do not test this compound trade-off
      reasoning.
    \item Games with multi-dimensional scoring are a natural testbed, but
      most such games are too complex to evaluate systematically
      (Civilization, Factorio) or too simple to differentiate models
      (multi-objective toy gridworlds).
  \end{itemize}
}

\plan{%
  What Yedi offers (contribution paragraph):
  \begin{itemize}
    \item A card game with 4 independent scoring dimensions (number, color,
      shape, word), each with multiple sub-modes. Merge outcomes are the
      PRODUCT of per-dimension multipliers, so one bad factor dominates.
    \item Dual resource constraints (mana budget + move budget) that force
      planning beyond greedy single-step optimization.
    \item A configurable difficulty surface: 40+ mode combinations from
      single-dimension (isolate one cognitive skill) to 4-dimension
      (compound trade-off reasoning).
    \item Two observation modes --- structured metadata and raw screenshots ---
      on the same underlying task, enabling controlled measurement of the
      perceptual grounding cost (``vision tax'').
    \item Open-source Gymnasium environment with WebSocket bridge to the
      real game engine, ensuring scoring fidelity.
  \end{itemize}
}

\plan{%
  Results preview (1 paragraph):
  \begin{itemize}
    \item Frontier LLMs beat a strong greedy baseline across all tested
      configs in metadata mode.
    \item Random agent defines a clear floor (score $\approx$ starting mana).
    \item Greedy agent provides a calibrated mid-point.
    \item Difficulty scales predictably with the number of active dimensions.
    \item \todo{Vision mode results once we have them.}
  \end{itemize}
}


% ======================================================================
\section{Related Work}
\label{sec:related}

Our work sits at the intersection of game-based AI evaluation, multi-objective
decision making, and vision-language model benchmarking. We organize prior work
into four areas.

\subsection{Game-Based Benchmarks for LLM and VLM Agents}

A growing body of work uses games to evaluate language and vision-language
models as interactive agents. SmartPlay~\citep{wu2024smartplay} was among the
first dedicated benchmarks, testing LLMs across six text-based games spanning
spatial reasoning, planning, and object dependencies. GameBench~\citep{costarelli2024gamebench}
extended this to nine game environments, each targeting a different strategic
reasoning axis, while GTBench~\citep{duan2024gtbench} introduced game-theoretic
structure with Monte Carlo Tree Search opponents across complete- and
incomplete-information settings. LLMArena~\citep{chen2024llmarena} and
GameArena~\citep{hu2025gamearena} further explored multi-agent competitive
evaluation and live human--AI interaction respectively.
DSGBench~\citep{tang2025dsgbench} shares our emphasis on multi-dimensional
evaluation, scoring agents across five specific capability dimensions within
complex strategy games.

On the vision side, lmgame-Bench~\citep{hu2026lmgamebench} wraps six
video/puzzle games in a unified Gymnasium API with perception scaffolds for
LLMs, while V-MAGE~\citep{zheng2025vmage} evaluates VLMs on arcade games
where all information is conveyed solely through screenshots.
Cradle~\citep{tan2025cradle} demonstrated general computer control via
screenshots in open-world games (Red Dead Redemption~2, Cities: Skylines),
and VideoGameBench~\citep{zhang2025videogamebench} tested VLMs on 23 classic
video games, finding that even the best model completes under 1\% of the
benchmark. GameWorld~\citep{ouyang2026gameworld} provides a browser-based
benchmark with state-verifiable scoring, comparing raw-input and
semantic-action interfaces.

Most directly relevant is BALROG~\citep{paglieri2025balrog}, which aggregates
six RL game environments (BabyAI, Crafter, TextWorld, MiniHack, NetHack) and
evaluates both LLMs and VLMs with text and image observations. Their central
finding---that several models perform \emph{worse} with visual
representations---motivates our controlled ``vision tax'' measurement. Our work
complements BALROG by focusing on a card game domain that tests
multi-dimensional strategic reasoning rather than grid-world navigation.

Earlier foundations include the Arcade Learning Environment~\citep{bellemare2013ale},
which established pixel-based game evaluation for RL; the NetHack Learning
Environment~\citep{kuttler2020nethack}, which supports multiple observation
modalities for the same underlying game; TextWorld~\citep{cote2018textworld}
for procedurally generated text adventures;
MineDojo~\citep{fan2022minedojo} and MineRL~\citep{guss2019minerl} for
Minecraft-based benchmarks; AgentBench~\citep{liu2024agentbench} for
multi-environment LLM agent evaluation; and
Atari-GPT~\citep{waytowich2024atarigpt}, which applied modern VLMs to classic
Atari games as zero-shot controllers.

\subsection{Card Game AI}

Card games pose distinct challenges for AI: imperfect information, stochastic
draws, and strategic depth within compact state spaces.
Pluribus~\citep{brown2019pluribus} achieved superhuman performance in six-player
no-limit Texas Hold'em using game-theoretic self-play, and
PokerBench~\citep{zhuang2025pokerbench} later evaluated LLMs on 11,000 poker
scenarios, finding that even GPT-4 achieves only 54\% accuracy.
The Hanabi Challenge~\citep{bard2020hanabi} proposed the cooperative card game
Hanabi as an AI benchmark emphasizing theory-of-mind reasoning under imperfect
information. More recently, \citet{wang2025llmcardgames} systematically
evaluated LLMs across eight diverse card games (DouDizhu, Mahjong, Uno, Gin
Rummy, Hold'em variants), finding that fine-tuning can approach strong baselines
but at the cost of general capability. RL approaches have also been applied to
complex collectible card games such as Hearthstone~\citep{hearthstone2023}.
Cicero~\citep{meta2022cicero} demonstrated human-level play in Diplomacy by
combining language models with strategic reasoning in a multi-player,
imperfect-information setting.

Yedi differs from these in two key respects: (1)~merge outcomes are determined
by the \emph{product} of per-dimension multipliers across 1--4 independent
cognitive dimensions, so one bad dimension ruins the compound gain; and (2)~the
benchmark supports both structured-metadata and screenshot observation modes for
the same underlying task, enabling controlled measurement of the perceptual
grounding cost.

\subsection{Multi-Objective Decision Making}

The multi-objective RL (MORL) literature addresses agents that must optimize
vector-valued rewards. \citet{hayes2022morl} provide a practical guide covering
utility functions, scalarization approaches, and Pareto-based methods for
sequential decision problems. Yedi's multiplicative scoring can be viewed as a
specific scalarization: $r = \prod_{d \in \text{active}} m_d$, where $m_d$ is
the per-dimension merge multiplier. Unlike additive scalarizations (where
strength in one dimension compensates for weakness in another), the
multiplicative form is dominated by the \emph{worst} factor---a property that
makes balanced competence across all dimensions essential.

This connects to resource-constrained planning in LLM agents.
AucArena~\citep{chen2024aucarena} tests LLMs on budget management and
adaptive strategy in auction settings, while \citet{huang2024planningsurvey}
survey LLM planning capabilities including task decomposition, plan selection,
and reflection---all of which Yedi requires within its dual-constraint
(mana budget + move budget) framework.

\subsection{Cognitive Benchmarks and Vision-Language Evaluation}

Standard LLM benchmarks---MMLU~\citep{hendrycks2021mmlu},
GPQA~\citep{rein2023gpqa}, MATH~\citep{hendrycks2021math},
BIG-bench~\citep{srivastava2023bigbench} and its hard
subset~\citep{suzgun2023bigbenchhard}---test knowledge and reasoning through
independent questions with additive scoring. A model scoring 90\% on each of
four subtasks achieves a 90\% average; a model achieving $0.9\times$ in each of
four Yedi dimensions scores $0.9^4 = 0.66\times$, dramatically penalizing any
weakness. Chollet's \emph{On the Measure of Intelligence}~\citep{chollet2019measure}
and the Abstraction and Reasoning Corpus (ARC) advocate for benchmarks that test
skill acquisition and generalization from cognitive-science priors. Yedi shares
this philosophy---each game configuration presents a novel combination of
dimension rules---but extends it to sequential decision-making under resource
constraints.

Recent work has begun applying psychometric frameworks to LLMs.
\citet{reddy2025chcllm} applied the Cattell--Horn--Carroll (CHC) theory of
cognitive abilities to frontier models, finding a ``catastrophic paradox''
where models excel at crystallized knowledge but fail psychometric scoring.
\citet{pellert2024aipsychometrics} demonstrated that standard psychometric
inventories can measure LLM psychological traits. Yedi can be viewed as a
\emph{behavioral} cognitive assessment that measures fluid reasoning (Gf),
quantitative reasoning (Gq), and visual processing (Gv) through observed
gameplay rather than self-report, potentially avoiding the validity concerns
these works raise.

For vision-language evaluation specifically, MMBench~\citep{liu2024mmbench},
MM-Vet~\citep{yu2024mmvet}, and SEED-Bench~\citep{li2024seedbench} test
fine-grained perceptual abilities in static settings.
SeeClick/ScreenSpot~\citep{cheng2024seeclick} and
ScreenQA~\citep{hsiao2025screenqa} focus on GUI grounding from screenshots.
OSWorld~\citep{xie2024osworld} and AndroidWorld~\citep{rawles2025androidworld}
test agents on real computer and mobile tasks, while
WebArena~\citep{zhou2024webarena} and
VisualWebArena~\citep{koh2024visualwebarena} provide web-based agent
benchmarks. VisualWebArena is particularly relevant as it compares agents
using accessibility trees (structured text) versus screenshots for the same
web tasks---directly analogous to our metadata-versus-vision design.
Our benchmark builds on the Gymnasium standard
interface~\citep{towers2024gymnasium} and, while Unity
ML-Agents~\citep{juliani2018unityml} provides RL integration for Unity games,
we use a WebSocket bridge to the WebGL build to enable browser-based deployment
and native screenshot capture.


% ======================================================================
\section{The Yedi Game}
\label{sec:game}

Yedi is a single-player card game originally developed as a cognitive-training
app on mobile platforms. Each card carries attributes across up to seven
independent cognitive dimensions --- four perceptual/reasoning dimensions
(\emph{number, color, shape, word}) and three secondary dimensions
(\emph{beat/music, memory, motor}). Players merge cards to grow their score
(``mana''), with each active dimension contributing an \emph{independent}
multiplier to the merge outcome. The multipliers compose multiplicatively
across dimensions, so optimizing the compound score requires balanced
competence across every active dimension: a single weak decision anywhere in
the cognitive stack collapses an otherwise strong merge. This section
describes the game mechanics that we inherit verbatim in the benchmark; the
benchmark-specific interface is deferred to Section~\ref{sec:benchmark}.
Readers who prefer to see the game played before reading its rules may
consult the video walkthroughs released by the original
developer.\footnote{\url{https://www.youtube.com/@mindfreegames7459}}

\subsection{Board and Actions}
\label{sec:board}

The board consists of one \emph{new} slot that receives freshly drawn cards,
five numbered \emph{build} slots that hold cards available for merging, and
one \emph{sell} slot that converts a card back into mana. The player issues
one of three action types per turn:

\begin{itemize}
  \item \textbf{Draw}: spawn a new card in the new slot at a fixed mana cost
    (10 mana per draw).
  \item \textbf{Move/Merge}: drag a card from its current slot to a target
    build slot. If the target is empty, the card is placed. If it is
    occupied, the two cards merge: the moving card is consumed and the
    target card's mana is multiplied by the product of per-dimension
    merge gains.
  \item \textbf{Sell}: drag a card to the sell slot. The card is destroyed
    and its current mana is refunded to the player.
\end{itemize}

Each card has a merge cap (3 merges at the easiest difficulty, rising to 5
at the hardest): once reached, the card must be sold to free the slot. The
player's score at any moment is the peak mana ever reached during the
episode, matching the scoring system used by the mobile app's leaderboards.

\subsection{Cognitive Dimensions}
\label{sec:dimensions}

Each dimension has multiple \emph{sub-modes} that change what constitutes a
good merge. Table~\ref{tab:dimensions} lists the four primary dimensions
used throughout our benchmark; the three secondary dimensions are
summarised below.

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Dimension} & \textbf{Sub-mode} & \textbf{Merge rule (informal)} \\
\midrule
\textbf{Number (Gq)} & add & $x + y = 0$ is perfect (number pair) \\
 & multiply & $x \cdot y = 1$ is perfect \\
 & gcd & $\gcd(x, y) > 1$ is perfect \\
 & trigon & trig identity combos \\
 & interval, vector, sort & sub-domain rules (see Appendix) \\
\midrule
\textbf{Color (Gv)} & add & RGB additive mixing to white is perfect \\
 & subtract & CMY subtractive mixing to black is perfect \\
 & gray & sum of colour indices $>9 \Rightarrow$ white \\
 & text & Stroop: text word must match fill colour \\
\midrule
\textbf{Shape (Gv/Gf)} & triangle, rectangle, triple & shape-rotation matching \\
 & sphere (11 shapes) & 3D orientation matching \\
 & kanizsa & illusory-contour completion \\
 & hanoi (17 shapes) & tower-of-Hanoi ordering (time-extended) \\
\midrule
\textbf{Word (Gc)} & verbs, nouns, adjectives, prepositions & antonym pairs \\
 & synVerbs, synAdjectives & synonym pairs \\
 & grammar, questions & grammatical/QA pairing \\
 & scrabble & concatenation forms a valid word \\
\bottomrule
\end{tabular}
\caption{Primary cognitive dimensions and their sub-modes. Each dimension
contributes an \emph{independent} multiplier to a merge; the merge outcome
is the product across active dimensions. Parenthetical labels map each
dimension loosely to Cattell--Horn--Carroll (CHC) cognitive
factors~\citep{reddy2025chcllm}.}
\label{tab:dimensions}
\end{table}

The three secondary dimensions add orthogonal pressures that can be stacked
on top of any primary-dimension configuration:

\begin{itemize}
  \item \textbf{Beat (Ga)}: merges gain a multiplier bonus or penalty
    depending on how close the drop is to the on-beat phase of a
    rhythmic pulse ($0.03$-wide ``perfect'' window, $0.5$-wide penalty
    band). Six beat patterns from double (1.8\,s period) to five
    (4.32\,s period).
  \item \textbf{Memory (Gsm)}: after a short visibility window, card face
    values are hidden behind a back sprite. The player must remember
    which slot held which card. Three sub-modes control the reveal
    policy (Section~\ref{sec:config}).
  \item \textbf{Motor (Gps)}: merges gain a tier bonus based on drag
    trajectory quality (speed, endpoint accuracy, how many intermediate
    slots were visited). Modelled in the benchmark via Fitts-law
    noise injection (Section~\ref{sec:env}).
\end{itemize}

\subsection{Merge Mechanics}
\label{sec:merge}

The multiplicative compound-reasoning bottleneck is the central game-design
decision that motivates Yedi as a benchmark. When two cards merge, each
active dimension $d$ independently produces a discrete \emph{gain tier}
$g_d \in \{-1, 0, 1, 2, 3\}$ evaluated from the dimension's merge rule.
Tiers map to multipliers via a difficulty-dependent table
(Table~\ref{tab:tiers}), and the surviving card's mana updates as

\begin{equation}
\label{eq:merge}
\text{mana}_{\text{new}} \;=\; \text{mana}_{\text{old}} \;\cdot\; \prod_{d \in \mathcal{A}} \mu(g_d)
\end{equation}

where $\mathcal{A}$ is the set of active dimensions for the configuration
and $\mu(\cdot)$ is the tier-to-multiplier map from Table~\ref{tab:tiers}.

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}clccc@{}}
\toprule
\textbf{Tier $g_d$} & \textbf{Label} & $\mu$ (Diff 0) & $\mu$ (Diff 1) & $\mu$ (Diff 2) \\
\midrule
$3$  & perfect  & $2.5$ & $2.0$ & $1.7$ \\
$2$  & great    & $2.0$ & $1.5$ & $1.3$ \\
$1$  & ok       & $1.5$ & $1.2$ & $1.1$ \\
$0$  & neutral  & $1.0$ & $1.0$ & $1.0$ \\
$-1$ & bad      & $0.9$ & $0.8$ & $0.7$ \\
\bottomrule
\end{tabular}
\caption{Per-dimension gain tier to multiplier. Higher in-game difficulty
compresses the payoff for perfect play and sharpens the penalty for bad
merges. The benchmark uses Diff~0 by default.}
\label{tab:tiers}
\end{table}

The consequence of the product form in Equation~\eqref{eq:merge} is that
\emph{no strong dimension compensates for a weak one}. Consider a 4-dim
configuration in which three dimensions merge at tier~3 (``perfect'') and
one at tier~$-1$ (``bad''). The compound multiplier is
$2.5 \cdot 2.5 \cdot 2.5 \cdot 0.9 \approx 14.06$, versus
$2.5^4 \approx 39.06$ for fully perfect play --- a $64\%$ loss driven
entirely by a single weak axis. In contrast, an additive scoring scheme
with the same per-dimension values gives $9.1$ vs.\ $10.0$, a $9\%$ loss.
The multiplicative form is the formal ground for Yedi's claim as a
multi-objective reasoning benchmark: agents that reason as if score were a
mean will under-optimise, and the gap widens with each additional active
dimension.

Table~\ref{tab:chain} makes the compounding explicit. A 3-merge chain that
holds tier~3 across all active dimensions grows mana from a nominal 10 to
over 600 in the 4-dim case; a single tier~$-1$ merge partway through the
chain cuts the final mana by more than half.

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Active dims} & \textbf{Tier~3 chain of 3} & \textbf{With one tier~$-1$} & \textbf{Ratio} \\
\midrule
1 dim & $10 \cdot 2.5^3 = 156.25$ & $10 \cdot 2.5^2 \cdot 0.9 = 56.25$ & $0.36$ \\
2 dim & $10 \cdot 2.5^6 = 2441.4$ & $10 \cdot 2.5^5 \cdot 0.9 \cdot 2.5 \approx 879$ & $0.36$ \\
3 dim & $10 \cdot 2.5^9 \approx 38{,}147$ & $\approx 13{,}733$ & $0.36$ \\
4 dim & $10 \cdot 2.5^{12} \approx 596{,}046$ & $\approx 214{,}577$ & $0.36$ \\
\bottomrule
\end{tabular}
\caption{Compound growth from a 3-merge chain. Column 2: all dimensions
tier~3 every merge. Column 3: one merge has a tier~$-1$ gain on one
dimension (others still tier~3). The per-merge hit is the same (a $0.36$
ratio); with 4 dimensions active it costs nearly $400{,}000$ mana.}
\label{tab:chain}
\end{table}

\todo{Figure: game-board screenshot with slot + dimension annotations.}


% ======================================================================
\section{Benchmark Design}
\label{sec:benchmark}

The benchmark wraps the Yedi game described in Section~\ref{sec:game} in a
Gymnasium-compatible environment that exposes the same merge rules, scoring
system, and card generators used by the mobile app. The full implementation
--- Unity WebGL build, Python environment, agent wrappers, benchmark
configurations, and analysis scripts --- is released as an open-source
package.\footnote{\url{https://github.com/yilmazozgur/yedi-v4}} Crucially, all scoring
logic remains in the Unity~C\# implementation; the Python side is a thin
control layer. This design trades a modest per-step overhead (WebSocket
round-trip) for complete fidelity to the reference implementation, so that
benchmark numbers are directly comparable with human leaderboard scores.

\subsection{Environment}
\label{sec:env}

The environment exposes a standard Gymnasium interface
(\texttt{reset()}/\texttt{step(action)}) whose step returns the observation,
a shaped reward, a \texttt{terminated} flag for game-over, a
\texttt{truncated} flag for the step cap, and an \texttt{info} dict with
full game state. A Python FastAPI process holds a WebSocket channel to a
headless Chromium that runs the Unity WebGL build; on each step the bridge
(i) freezes game time, (ii) submits the action via an injected JavaScript
hook, (iii) awaits the Unity-side callback with the updated state, and
(iv) resumes game time. Pausing between actions removes wall-clock
pressure, so model latency never confounds reasoning quality (motor noise
and beat timing are re-introduced explicitly as ablation knobs below).

The action space is \texttt{Discrete}(37): action $0$ draws a card into the
new slot; actions $1\ldots30$ encode a \textsc{move/merge} from source slot
$s \in \{\text{new}, 1, 2, 3, 4, 5\}$ to destination slot
$d \in \{1, 2, 3, 4, 5\}$ as $a = 1 + 5s + (d{-}1)$; actions $31\ldots36$
encode a \textsc{sell} from source slot $s$. Every observation carries a
\texttt{MultiBinary(37)} \emph{action mask} derived from current slot
occupancy, merge caps, and game-over state, so agents never need to reason
about low-level legality. Episodes truncate at a configurable step cap
(default $100$), replacing the mobile app's wall-clock timer; termination
occurs when a draw would drop mana below zero. Random seeding is exposed end
to end: the seed controls card spawn order and any stochastic merge effects,
so every run is replayable.

\subsection{Observation Modes}
\label{sec:obs}

The same game state is rendered in one of two channels:
\begin{itemize}
  \item \textbf{Metadata.} A structured dump of the seven slots with each
    card's active-dimension attributes (e.g.\ \texttt{Num=3, Color=red,
    Shape=triangle-up, Word=help}), current and maximum mana, merge counter
    and cap, the list of active dimensions and their sub-modes, the step
    counter, and the action mask. All information required to reason about
    the next merge is explicit; nothing is hidden except state the rules
    hide (the Memory dimension).
  \item \textbf{Vision.} A $960{\times}540$ PNG screenshot of the live
    canvas alongside a minimal HUD (current mana, step count, action mask).
    Card attributes are \emph{not} printed; the model must read them from
    pixels. The board layout, slot positions, and rendering pipeline match
    the human mobile-app experience.
\end{itemize}

Orthogonal to the observation channel, the benchmark offers three
\emph{history modes} that control what an agent sees across successive
steps: \textsc{stateless} (each step is a fresh prompt, no prior context),
\textsc{conversational} (the episode accumulates as a rolling
user-assistant chat), and \textsc{compact} (a terminal-style transcript
that appends one fixed-width line per completed step summarising the action
and resulting slot state). Compact mode is designed to give long-horizon
reasoning to LLMs whose cost or context budget precludes a full
conversational history; it also lets us decouple \emph{memory} from
\emph{dialogue context} when analysing failure modes. All three modes are
available for metadata and vision observations, yielding six $($observation,
history$)$ combinations that a single config can be swept across.

\subsection{Configurations}
\label{sec:config}

A \emph{configuration} fixes which dimensions are active and which sub-mode
each dimension uses (e.g.\ \texttt{\{number:add, color:add\}}). We provide a
curated library of 36 configurations grouped into four difficulty tiers
(Table~\ref{tab:config-summary}); the full list is catalogued in
Appendix~\ref{app:configs}.

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}lrll@{}}
\toprule
\textbf{Tier} & \textbf{\# configs} & \textbf{Dimensions active} & \textbf{Sub-modes used} \\
\midrule
Easy      & 8  & 1                     & simplest per-dimension (add, triangle, verbs, \ldots) \\
Medium    & 10 & 1 hard mode, or 2 & gcd, gray, kanizsa, grammar; first 2-D and memory combos \\
Hard      & 10 & 2--3 with hard modes  & Stroop, tower-of-Hanoi, scrabble, 3-D add combos \\
BrainFry  & 8  & 3--5 hard modes       & the Yedi signature 4-D, memory + 4-D, all-hard 5-D \\
\bottomrule
\end{tabular}
\caption{Configuration tiers. Each tier escalates along two axes: the
number of active dimensions, and the difficulty of the per-dimension
sub-mode. Easy isolates a single cognitive skill; BrainFry stresses
compound reasoning with up to five interacting dimensions.}
\label{tab:config-summary}
\end{table}

The tiering is deliberate rather than empirical: Easy configs give a clean
per-dimension readout where metadata agents \emph{should} excel and vision
agents are stress-tested on a single axis of perception. Medium introduces
the first 2-D products and the first memory-combined configs. Hard
introduces the Stroop (color:text) and tower-of-Hanoi (shape:hanoi)
sub-modes that are designed to break VLMs, as well as 3-D products.
BrainFry is the regime in which \emph{every} frontier model should be
measurably imperfect: classic 4-D (number+color+shape+word at their easiest
sub-modes), the same 4-D with every-action memory hiding, and 5-D
all-hard.

\subsection{Metrics}
\label{sec:metrics}

Our primary metric is the \emph{max mana} achieved during an episode: the
single scalar shown on the human leaderboard and the quantity whose
multiplicative growth the merge mechanics are designed around. Because mana
compounds multiplicatively, we report its distribution on a log scale.

We additionally track: the \emph{mana surplus}
(\,$\max\_\text{mana} - \text{starting\_mana}$\,, which equals zero for an
agent that only drains and never merges productively); the distribution of
per-merge gain tiers (what fraction of merges were \textsc{perfect},
\textsc{great}, \textsc{ok}, \textsc{neutral}, \textsc{bad}); the mean
number of merges completed on cards that were eventually sold
(\emph{sell timing}, a proxy for planning depth); the number of wasted
draws (draws that ended without a profitable merge); and, when merge
previews are visible, the \emph{preview adherence rate} --- how often the
agent picks the merge the greedy baseline would have picked.

To aggregate across configs with very different mana scales, we map each
episode's max mana through a logistic percentile calibrated against a
human reference distribution (Appendix~\ref{app:percentile}), yielding a
$0$--$100$ score that is comparable across tiers and configs and stackable
into radar charts.

\subsection{Ablation Knobs}
\label{sec:ablations}

Several aspects of the environment that are fixed in human play are
independently toggleable in the benchmark. We use these to disentangle
\emph{what} a model struggles with from \emph{that} it struggles:

\begin{itemize}
  \item \textbf{Merge preview}: when enabled, every observation includes,
    for each legal merge, the exact tier-$g_d$ that each active dimension
    would produce. This exposes whether an agent's bottleneck is
    \emph{evaluating} candidate merges or \emph{planning} among evaluated
    merges.
  \item \textbf{Perfect memory}: when active, cards are never hidden even
    under memory sub-modes. Toggling against the default memory hiding
    isolates short-term memory capacity from strategy.
  \item \textbf{Motor skill} (Fitts-law noise $\sigma$): sets the
    variability of the drag endpoint used to evaluate motor-dimension
    merges. A benchmark-wide default ensures fairness; per-agent overrides
    are logged into the episode trace.
  \item \textbf{History mode}: \textsc{stateless} vs.\ \textsc{compact} vs.\
    \textsc{conversational}, per Section~\ref{sec:obs}.
\end{itemize}

Any subset of these may be frozen to isolate a single causal axis; e.g.\
the vision-tax experiment freezes merge preview \emph{off} and history mode
to \textsc{compact}, varying only the observation channel.

\subsection{Fairness Invariants}
\label{sec:fairness}

Three invariants are maintained across all agents and configs. First, game
time is frozen between actions, so latency cannot influence score; the
Motor and Beat dimensions re-introduce time sensitivity \emph{only} when
explicitly enabled. Second, memory hiding is \emph{turn-based}, not
wall-clock-based: a hidden card becomes visible for exactly one agent step
after being touched, then re-hides, regardless of how long the model
deliberated. Third, card generation is deterministically seeded so that
every $(\text{config}, \text{seed})$ pair produces the same card stream
and identical merge outcomes; this lets us compare agents on matched
episodes rather than matched distributions.


% ======================================================================
\section{Agents}
\label{sec:agents}

We evaluate four agent families. \textsc{Random} and \textsc{Greedy} are
programmatic baselines that bracket the range of scores achievable
\emph{without} strategic reasoning: Random establishes the floor, and
Greedy establishes the ceiling reachable from perfect local information
with no planning. \textsc{LLM} and \textsc{VLM} are model agents that
consume the metadata and vision observation channels respectively and
return actions through the same Gymnasium interface. All four use the
shared \texttt{BaseAgent} protocol, so an experiment sweep is a single
$(\text{agent}, \text{config}, \text{seed})$ cross-product.

\subsection{Random Agent (Floor)}
\label{sec:random-agent}

The random agent samples uniformly from the valid actions indicated by the
observation's action mask. It issues no draws unless drawing is the only
legal action, merges and sells at the same rate, and never considers
merge quality. Over a typical episode its per-merge gain tier distribution
is centred on $g_d = 0$, so merges neither grow nor shrink mana in
expectation; draws, by contrast, deterministically cost 10 mana each.
Without a preferred direction the trajectory drifts down, and the expected
max mana stays within noise of the starting mana (200). The random agent
therefore serves two purposes: it is the score floor against which to
measure progress, and it is the sanity check that the environment is
actually producing useful learning signal --- an agent that cannot beat
random is not reasoning about the game.

\subsection{Greedy Agent (Calibrated Baseline)}
\label{sec:greedy-agent}

The greedy agent reads the merge-preview table that the WebSocket bridge
ships with every observation, which contains, for each legal merge, the
per-dimension gain tier $g_d$ computed by the same C\# code that scores
the real game. At each step it follows a fixed checklist that mirrors the
decision order exposed in the LLM system prompt:
\begin{enumerate}
  \item If any build slot has hit its merge cap, sell that slot.
  \item If the new slot is occupied, score every legal
    new $\to$ slot$_n$ preview. If the best candidate is at least
    \emph{ok} ($g_d \geq 1$) on \emph{every} active dimension, execute
    that merge.
  \item Otherwise, if there is an empty build slot, place the new card
    there. Otherwise, sell the new card rather than corrupt a built slot.
  \item If the new slot is empty, draw.
\end{enumerate}
The scoring rule is the highest \emph{minimum} per-dimension tier rather
than sum or mean, because the merge outcome compounds multiplicatively
(Equation~\ref{eq:merge}): a single weak dimension dominates the loss.
Greedy does no lookahead beyond one step and does not consider inter-slot
merges (slot$_a \to$ slot$_b$), even though the bridge ships those previews
too; this keeps it an \emph{information} baseline rather than a
\emph{planning} baseline. The gap between Greedy and an LLM at the same
config therefore measures the value of planning given already-perfect
local information, while the gap between Greedy and Random measures the
value of local information given no planning.

\subsection{LLM Agent (Metadata Mode)}
\label{sec:llm-agent}

The LLM agent consumes the metadata observation and returns a single
integer action per step. The prompt is a fixed three-part template: a
system message that states the game rules, the merge-gain tiers, and the
decision checklist from Section~\ref{sec:greedy-agent}; a game-state block
that renders the current observation as structured text (active
dimensions and sub-modes, per-slot card attributes, mana, step counter,
legal-action list); and, in history-enabled modes, a trail of prior
$(\text{action}, \text{outcome})$ rows. The agent's output is constrained
to the legal-action list, so malformed or off-manifest responses become
no-ops that the environment drops without advancing game state.

All three history modes from Section~\ref{sec:obs} are supported.
\textsc{Stateless} issues one independent API call per step, which costs
the most tokens per decision but eliminates context drift.
\textsc{Conversational} accumulates a rolling user--assistant transcript
within an episode, letting the model reference earlier reasoning but
paying quadratic cost in history length. \textsc{Compact} appends one
fixed-width line per completed step --- a terminal-style transcript of
slot states and merge tiers --- which lets us give a $100$-step horizon to
models whose context or price budget would not tolerate the equivalent
conversation. The merge-preview field is an independent toggle: when
enabled, the per-dimension tier table appears alongside the slot dump and
the model only has to plan, not evaluate.

We evaluate a range of commercial and open-weight models spanning
parameter scale and release family. \todo{Confirm final model list from
the UI-visible run manifest before publication.}

\subsection{VLM Agent (Vision Mode)}
\label{sec:vlm-agent}

The VLM agent consumes the vision observation --- a $960{\times}540$ PNG
screenshot of the live canvas --- alongside a minimal HUD (current mana,
step counter, legal actions). Card attributes are \emph{not} enumerated in
text: the model must read number, colour, shape, and word off the pixels.
The system prompt, decision checklist, and action-encoding rules are
identical to the metadata agent, so the only source of score differences
between a model's metadata run and its vision run is visual grounding,
not strategy. All three history modes are supported; in
\textsc{compact}~+~vision mode, trail entries still list actions and
mana, but slot attributes are blanked so the screenshot remains the sole
source of perceptual information.

Running the same config with the same model in both channels yields a
\emph{vision tax}: the multiplicative score ratio between metadata and
vision, or equivalently the additive drop in the logistic percentile
(Section~\ref{sec:metrics}). We report it per-dimension by restricting to
Easy configs whose sole active dimension is perceptual (colour, shape).
\todo{Confirm final VLM list once the UI-visible run manifest is
finalised; the current Kimi-K2.5 runs abandoned before episode end and
must be rerun.}

\subsection{Human Baseline}
\label{sec:human-agent}

For reference we include a small number of human sessions collected
through the benchmark UI's recorder, bound to the same Gymnasium
interface and scoring as the model agents. These are used only to
calibrate the logistic percentile (Appendix~\ref{app:percentile}) and to
anchor the Easy-tier comparison; no human experiments are involved in
the quantitative claims of Section~\ref{sec:experiments}.


% ======================================================================
\section{Experiments}
\label{sec:experiments}

\plan{%
  This section will be filled in as we run more benchmarks.
  Structure the results around 3 research questions.
}

\subsection{RQ1: Do LLMs exhibit multi-dimensional strategic reasoning?}
\plan{%
  \begin{itemize}
    \item Compare LLM agents vs.\ greedy vs.\ random across single-dim
      and multi-dim configs.
    \item Hypothesis: LLMs should beat greedy on multi-dim configs (where
      planning and trade-off reasoning matter) even if greedy has perfect
      local merge scores.
    \item Metric: max mana, broken down by number of active dimensions.
    \item \todo{Run matrix: 3+ LLMs $\times$ 8+ configs $\times$ 5 episodes.}
  \end{itemize}
  \todo{Table: agent $\times$ config score matrix.}\\
  \todo{Figure: bar chart grouped by dimension count.}
}

\subsection{RQ2: What is the vision tax?}
\plan{%
  \begin{itemize}
    \item Run the same LLM in metadata mode and vision mode on identical
      configs.
    \item The score difference isolates perceptual grounding cost from
      strategic reasoning.
    \item Break down by dimension type: color (requires seeing actual RGB),
      shape (spatial), number (OCR), word (text reading).
    \item Hypothesis: vision tax is smallest for number/word (essentially
      OCR) and largest for color/shape (genuine visual understanding).
    \item \todo{Run matrix: 2+ VLMs $\times$ same configs $\times$
      metadata and vision $\times$ 5 episodes.}
  \end{itemize}
  \todo{Table: metadata score vs.\ vision score, with delta column.}\\
  \todo{Figure: paired comparison per config.}
}

\subsection{RQ3: How does performance scale with model capability?}
\plan{%
  \begin{itemize}
    \item Test models across a capability range: small open-source
      ($\sim$7B), medium ($\sim$70B), large ($\sim$120B+), frontier
      (Claude, GPT-4).
    \item Does Yedi score correlate with general benchmarks (MMLU, HumanEval)?
      Or does it measure something orthogonal?
    \item Hypothesis: multi-dim configs should differentiate models more
      than single-dim configs (harder task = more headroom).
    \item \todo{Run matrix: 5+ models $\times$ easy/medium/hard configs
      $\times$ 5 episodes.}
  \end{itemize}
  \todo{Figure: scaling curve (model capability vs.\ Yedi score).}\\
  \todo{Scatter: Yedi score vs.\ MMLU / HumanEval.}
}

\subsection{Ablation: Merge Previews}
\plan{%
  \begin{itemize}
    \item With previews: agent receives pre-computed merge quality scores.
      Tests pure planning (``given that you know the quality of every
      possible merge, can you pick the right sequence?'').
    \item Without previews: agent must compute merge quality from card
      attributes and game rules. Tests evaluation + planning.
    \item Delta measures the ``evaluation bottleneck'' --- how much
      performance is lost because the model can't accurately score merges.
    \item \todo{Run same model with and without previews on same configs.}
  \end{itemize}
}


% ======================================================================
\section{Analysis}
\label{sec:analysis}

\plan{%
  Deeper dives beyond the headline numbers.
}

\subsection{Merge Quality Distribution}
\plan{%
  \begin{itemize}
    \item For each agent, plot the distribution of merge tiers chosen
      (PERFECT, GREAT, GOOD, BAD).
    \item Do LLMs make fewer BAD merges than greedy? (Greedy has perfect
      local info but no planning; LLMs might avoid traps greedy walks into.)
    \item Mana trajectory plots: how does mana evolve over the 100-step
      episode for each agent type?
  \end{itemize}
  \todo{Figure: merge tier histograms per agent.}\\
  \todo{Figure: mana trajectory curves (mean $\pm$ std).}
}

\subsection{Failure Modes}
\plan{%
  \begin{itemize}
    \item Qualitative analysis of LLM mistakes from trace logs.
    \item Common failure patterns: single-slot fixation, forgetting to sell,
      bad-merge spam, ignoring one dimension.
    \item Do different models exhibit different failure modes?
    \item \todo{Pull concrete examples from trace.jsonl files.}
  \end{itemize}
}

\subsection{Dimension-Specific Performance}
\plan{%
  \begin{itemize}
    \item Per-dimension merge quality breakdown: is the model better at
      number reasoning than color reasoning?
    \item Radar chart: one axis per dimension, comparing agents.
    \item Which dimension is the bottleneck for each model?
  \end{itemize}
  \todo{Figure: radar chart comparing agents across dimensions.}
}


% ======================================================================
\section{Multi-Agent Yedi: Adversarial and Cooperative Play}
\label{sec:multi-agent}

The single-agent benchmark of Sections~\ref{sec:benchmark}--\ref{sec:experiments}
measures an agent's ability to reason under multi-dimensional trade-offs in
isolation. We now describe a natural extension --- \emph{Multi-Agent Yedi}
--- in which two (or more) agents share a board, alternate actions, and each
accumulate their own private score. The design is already supported by the
WebSocket bridge; what changes is the scoring model in the Unity-side
C\# implementation, the per-agent prompt construction, and the research
questions the benchmark can now answer. We present the design in full
because it is, to our knowledge, the first \emph{mixed-motive strategic
reasoning} benchmark for frontier LLMs: previous multi-agent benchmarks
for language models are either fully cooperative
(Hanabi~\citep{bard2020hanabi}) or fully adversarial (poker and game-theory
suites~\citep{brown2019pluribus, duan2024gtbench, zhuang2025pokerbench,
chen2024aucarena}), while Diplomacy-style work~\citep{meta2022cicero,
mukobi2024welfare} embeds negotiation in a game whose board state is too
large to be used as a clean reasoning probe.

\subsection{Motivation}
\label{sec:ma-motivation}

The mixed-motive regime --- where agents must \emph{simultaneously}
cooperate to grow a shared pie and compete to claim slices of it --- is
the empty corner of the LLM-game-benchmark landscape. Melting
Pot~\citep{leibo2021meltingpot} established the importance of this regime
for reinforcement-learning agents, demonstrating that purely cooperative
or purely adversarial benchmarks do not predict behaviour in real-world
multi-party settings. Recent work on LLMs in repeated social
dilemmas~\citep{akata2023playing, lore2023strategic} finds that even
frontier models exhibit brittle strategic reasoning: they cooperate in
semantic framings that evoke cooperative scripts and defect in framings
that evoke competitive scripts, even when the underlying payoff matrix is
identical. What is missing is a benchmark that (i) presents a novel
game whose payoff structure is not already encoded in training data,
(ii) produces continuous-valued outcomes rather than a binary
cooperate/defect, and (iii) varies the cooperation--competition balance
along a continuous configuration axis. Multi-Agent Yedi meets all three:
its multiplicative scoring makes every merge either pie-growing or
pie-taking; its 40+ configurations parametrise how much of the joint
surplus can be privately captured; and its mechanics are sufficiently
idiosyncratic that no model has seen training transcripts of expert play.

\subsection{Game Rules for Two-Agent Yedi}
\label{sec:ma-rules}

Agents $A$ and $B$ alternate turns on a single shared board. Each turn the
active agent takes exactly one action from the same 37-action space as
the single-player game (Section~\ref{sec:env}) with the same action mask.
The shared state comprises one new slot, five build slots, one sell slot,
one board-level mana pool, and one shared card deck seeded deterministically
as in the single-player benchmark.

\paragraph{Mana accounting.} A single board-level mana variable
$M_\text{board}$ replaces the single-player mana. In addition, each agent
has a private score variable $M_A, M_B$. On a merge executed by agent
$X \in \{A, B\}$, let $\Delta = M_\text{board}' - M_\text{board}$ be the
net change in board mana driven by the merge (positive for productive
merges, negative for failed merges). Then:
\begin{equation}
\label{eq:ma-score}
M_X \mathrel{+}= \Delta, \qquad M_\text{board} \mathrel{+}= \Delta.
\end{equation}
The active agent is credited with the full delta of the move they made.
Draws cost $10$ mana, debited from $M_\text{board}$ only; the agent who
drew pays nothing from their private score. Sells refund the card's
current mana to $M_\text{board}$ only, leaving private scores unchanged.
Termination occurs when $M_\text{board}$ would go below zero on a draw
(identical to the single-agent rule); the episode's final scores are
$(M_A^{\text{peak}}, M_B^{\text{peak}})$ at the moment $M_\text{board}$
peaked.

\paragraph{Why this specific attribution.} The active-agent-captures-delta
rule is the minimal change that (i) preserves the meaning of the primary
metric (max mana $\to$ peak private mana) and (ii) creates a non-trivial
mixed-motive game. If instead both agents received the full delta, the
game reduces to fully cooperative play with turn-taking, and there is no
incentive to deny. If only the executing agent received the delta and
$M_\text{board}$ did not track it, there would be no shared resource and
the game reduces to two parallel single-agent games with interference
only through slot occupancy. Equation~\ref{eq:ma-score} is the unique
choice in which every merge simultaneously feeds the shared pool
\emph{and} the private score, so each action forces a choice between
``build now so I can harvest later'' and ``harvest now before they
do.''

\paragraph{The last-merge phenomenon.} Because mana compounds
multiplicatively (Equation~\ref{eq:merge}), the \emph{final} merge on a
slot --- the one that closes it at its merge cap of 3--5 --- carries
enormously more mana delta than any earlier merge on that slot. A slot
that has been built up to mana $m$ with one merge remaining yields
$m \cdot (\mu^* - 1)$ to whoever lands the final merge, where
$\mu^*$ is the compound multiplier for that final move. The agent who
\emph{built} the slot invested in capital that the agent who
\emph{closes} the slot harvests. This asymmetry is the engine of the
game's strategic richness: it simultaneously creates an incentive to
build (the board pool grows), an incentive to race (final merges are
disproportionately valuable), and an incentive to deny (selling a
built-up slot on your turn aborts the opponent's upcoming harvest).

\subsection{Strategic Primitives}
\label{sec:ma-primitives}

The two-agent game admits a taxonomy of move types that does not exist in
the single-player game. We enumerate them to give the reader a concrete
sense of what ``good play'' means in this regime.

\begin{itemize}
  \item \textbf{Selfish merge}: the active agent executes a merge
    with high expected $\Delta$ on a slot they can close themselves in
    a future turn. Zero-sum vis-\`a-vis the opponent; the safest
    move.
  \item \textbf{Setup / gift merge}: the active agent places a card in a
    build slot or executes a low-$\Delta$ merge that raises a slot's
    mana. Positive for $M_\text{board}$, but sets up a harvest whose
    harvester is undetermined. Rational only if the agent expects to
    recapture the harvest, or if the opponent is trusted not to.
  \item \textbf{Denial sell}: the active agent sells a slot the opponent
    is plausibly about to close at high $\Delta$. Refunds a small
    amount of mana to $M_\text{board}$ (no private-score change) at the
    cost of losing the slot entirely. Defects against any implicit
    cooperation.
  \item \textbf{Handoff merge}: a merge whose merger was made possible
    only by the opponent's prior setup. Equivalent in mechanics to a
    selfish merge, but socially coded as ``taking.'' Repeated handoffs
    without reciprocation are the multi-agent Yedi analogue of
    prisoner's-dilemma defection.
  \item \textbf{Draw refusal}: skipping the draw (e.g.\ by merging or
    selling when the new slot is empty) to force the opponent to pay
    the $10$-mana draw cost. Small magnitude but measurable over an
    episode.
  \item \textbf{Reciprocal alternation}: a stable cooperation policy in
    which each agent builds on odd turns and harvests on even (or
    vice versa), matching the turn cadence so that the partition of
    harvests is approximately equal. We expect frontier LLMs to reach
    this pattern in communication-enabled symmetric games and to
    \emph{fail} to reach it in no-communication games.
\end{itemize}

\subsection{Configuration Dimensions}
\label{sec:ma-configs}

Multi-Agent Yedi adds four orthogonal configuration axes to the
single-player surface:

\begin{enumerate}
  \item \textbf{Symmetry}: whether the two agents share the same active
    dimensions and sub-modes (symmetric) or see different active
    dimensions (asymmetric, e.g.\ $A$ plays \texttt{number:add} while
    $B$ plays \texttt{color:add} on the same cards). Asymmetric configs
    test whether a model can reason about an opponent whose scoring
    rule it does not share.
  \item \textbf{Communication channel}: no channel (pure observation-only
    equilibrium); a single-turn natural-language message slot (one
    message from each agent, before any action); or a free-form
    chat channel interleaved with actions. This is the bandwidth
    parameter of Welfare Diplomacy~\citep{mukobi2024welfare} applied to
    a game with a mechanical score.
  \item \textbf{Turn order}: fixed-alternation ($A, B, A, B, \ldots$);
    random-order (a fair coin flip each turn); or bid-for-priority
    (each agent privately submits a mana bid; winner acts, pays bid to
    the pool). Bid-for-priority is a stress test for models that must
    reason about their opponent's willingness to pay.
  \item \textbf{Opponent visibility}: whether each agent can see the
    opponent's cumulative private score during play, only at episode
    end, or never. Hidden-score variants probe whether models can
    infer an opponent's standing from action patterns alone.
\end{enumerate}

The Cartesian product of these axes with the single-player 36
configurations yields several thousand distinct multi-agent benchmarks.
We propose a curated initial suite (Appendix~\ref{app:ma-configs})
focused on symmetric same-dimension games across the four tiers, plus
three asymmetric probes at Medium, and a Welfare-Diplomacy-style
cooperation ladder (no-comm $\to$ one-shot comm $\to$ free chat) on a
single Easy config as the primary cooperation-capacity benchmark.

\subsection{Metrics}
\label{sec:ma-metrics}

Every single-agent metric (Section~\ref{sec:metrics}) retains its meaning
when computed per-agent. We add four multi-agent-specific scalars:

\begin{itemize}
  \item \textbf{Joint welfare} $W = M_A^{\text{peak}} + M_B^{\text{peak}}$:
    how much total private score was generated. Cooperation grows $W$;
    denial shrinks it. Directly comparable across symmetric configs.
  \item \textbf{Inequality} $I = |M_A^{\text{peak}} - M_B^{\text{peak}}|
    / W$: how asymmetrically the private scores were partitioned. An
    $I$ of $0$ is perfect parity; $I$ of $1$ is winner-take-all. Used to
    distinguish purely-cooperative outcomes ($W$ large, $I$ small) from
    cooperate-then-betray ($W$ large, $I$ large).
  \item \textbf{Gift ratio}: the fraction of the mana an agent added to
    $M_\text{board}$ via setup merges that was subsequently captured by
    the opponent rather than the setter. High gift ratio with high $W$
    indicates stable turn-taking; high gift ratio with low $W$
    indicates exploited cooperation.
  \item \textbf{Denial rate}: the fraction of the opponent's built-up
    slots that were sold by the active agent before the opponent could
    close them. Primary signal of defection.
\end{itemize}

In addition, when the communication channel is enabled, we log every
utterance and score (a) \textbf{promise fulfilment rate}: of moves
committed to in chat, the fraction the agent actually executed, and
(b) \textbf{opponent-model predictive accuracy}: a held-out agent
(``oracle'') reads the chat transcript and predicts the opponent's next
action; accuracy measures how much information the chat conveys.

\subsection{Research Questions}
\label{sec:ma-rqs}

Multi-Agent Yedi enables a distinct class of research questions
unreachable with the single-player benchmark:

\begin{itemize}
  \item \textbf{MA-RQ1 (Zero-shot cooperation).} Do frontier LLMs reach
    the reciprocal-alternation equilibrium on a symmetric game with no
    communication channel? The hypothesis, motivated by
    \citet{akata2023playing}, is \emph{no}: agents default to selfish
    merging and only cooperate when the prompt semantically frames the
    game as cooperative. Multi-Agent Yedi tests this without a framing
    confound because the game is novel.
  \item \textbf{MA-RQ2 (Communication value).} How much of the
    single-agent $\to$ symmetric-joint welfare gap is closed by
    enabling a one-shot natural-language message? How much more by
    free chat? We expect diminishing returns for frontier models and
    inverted returns for weaker models (more chat = more hallucinated
    commitments).
  \item \textbf{MA-RQ3 (Sabotage detection).} When $A$'s policy is
    scripted to defect (e.g.\ always deny-sell), does $B$ detect this
    within an episode and switch from cooperate to retaliate? Measured
    by the trajectory of $B$'s gift ratio over the episode.
  \item \textbf{MA-RQ4 (Asymmetric-capability carry).} When a strong
    model plays a weak model, does the strong model adapt to carry the
    weak partner (maximising joint welfare) or exploit it (maximising
    own score)? We expect frontier models to carry on cooperative-
    framed prompts and exploit on neutral prompts.
  \item \textbf{MA-RQ5 (Compositional scoring rules).} On asymmetric
    configs where $A$ scores \texttt{number} and $B$ scores
    \texttt{color}, does either agent learn the opponent's scoring
    rule by observing the opponent's preferences across merges? This
    is a clean test of inverse reinforcement learning in context.
\end{itemize}

\subsection{Technical Implementation Plan}
\label{sec:ma-impl}

The existing WebSocket bridge already serialises and deserialises full
game state per step. Four changes are required:

\begin{enumerate}
  \item \emph{C\# scorer split}: \texttt{ScoreManager} currently writes
    every merge delta into the single \texttt{currentMana} field. We
    add a \texttt{currentActingAgent} field set by the bridge before each
    step, and a dictionary of per-agent peak scores updated in the same
    place the current peak is updated.
  \item \emph{Bridge protocol}: the \texttt{action} command grows an
    \texttt{agent\_id} field; the \texttt{observation} payload grows
    per-agent score fields and the \texttt{current\_acting\_agent}
    field. Everything else (action space, mask, observation schema) is
    unchanged.
  \item \emph{Python orchestrator}: a new \texttt{MultiAgentEnv} wraps
    the existing \texttt{YediEnv} and exposes a \texttt{step(agent\_id,
    action)} interface. A turn-order policy (fixed / random / bid)
    decides which agent's \texttt{step} call is accepted next; others
    return a no-op.
  \item \emph{Prompt plumbing}: the LLM agent's system prompt gains an
    \textsc{opponent} block describing the opponent's active dimensions
    (for asymmetric configs) and the current chat transcript (when
    enabled). Each agent's prompt is independently assembled.
\end{enumerate}

We estimate the end-to-end implementation at 1--2 weeks of engineering,
primarily in the C\# scorer split and the Python orchestrator; the
prompt plumbing is a thin extension of the existing LLM agent code.

\subsection{Scaling Beyond Two Agents}
\label{sec:ma-nagents}

The attribution rule in Equation~\ref{eq:ma-score} generalises to $n$
agents with no modification: the active agent captures the full delta on
each action; all other agents observe. Three $\geq 3$-agent game modes
become interesting:

\begin{itemize}
  \item \textbf{Round-robin} ($n$ agents, fixed turn order): the
    natural generalisation. Reciprocal-alternation patterns become
    harder to sustain as $n$ grows because each agent has a smaller
    share of the turn clock.
  \item \textbf{Coalition formation}: a pre-action phase allows
    subsets of agents to publicly announce a merger of their private
    scores. Post-merge, $M_C = \sum_{i \in C} M_i$ is treated as a
    single score; the coalition plays as a meta-agent with an internal
    protocol for which agent acts on which turn. Tests whether LLMs
    can commit to coalitions and defect stably.
  \item \textbf{Tournament play}: $n$ agents play a series of two-agent
    matches in a round-robin, with leaderboard ranking by total joint
    welfare earned across all matches. The LLM arena
    paradigm~\citep{chen2024llmarena} meets continuous scoring.
\end{itemize}

We leave empirical investigation of the $n > 2$ regimes to future work
but note that the $n = 2$ case already admits several thousand
benchmark configurations and produces the strongest signal per episode,
justifying it as the primary multi-agent Yedi modality.


% ======================================================================
\section{Limitations}
\label{sec:limitations}

\plan{%
  Be upfront. Reviewers will find these anyway.
  \begin{itemize}
    \item \textbf{Visual complexity}: the game UI is clean and structured.
      Vision mode tests ``can you read a game board'' more than ``can you
      understand a complex visual scene.'' The vision tax may underestimate
      VLM limitations on messier real-world visual reasoning.
    \item \textbf{Game-specific knowledge}: models that have encountered
      Yedi-like rules in training data may have an advantage. We mitigate
      this with the diversity of mode combinations and the novelty of the
      specific merge mechanics, but cannot fully rule it out.
    \item \textbf{Bridge fragility}: the WebSocket bridge to Unity WebGL
      introduces engineering complexity and occasional disconnections.
      Episode-level error recovery is implemented but adds noise.
    \item \textbf{Evaluation cost}: each LLM episode requires $\sim$100
      API calls. Large-scale evaluation across many models and configs is
      expensive. We report results with limited episode counts and
      acknowledge the statistical uncertainty.
    \item \textbf{Single-player}: Yedi is not adversarial. The benchmark
      tests optimization against a fixed environment, not adaptation to
      an opponent. Strategic depth is bounded by the merge mechanics
      rather than an open-ended opponent.
    \item \textbf{Action space simplicity}: 37 discrete actions is not a
      complex action space. The challenge is in state evaluation and
      planning, not in action selection or motor control.
  \end{itemize}
}


% ======================================================================
\section{Conclusion}
\label{sec:conclusion}

\plan{%
  3 paragraphs:
  \begin{enumerate}
    \item We presented Yedi, a benchmark for multi-dimensional strategic
      reasoning. The multiplicative interaction across dimensions means
      models must optimize the worst-case factor, not the average ---
      a capability not tested by existing single-axis game benchmarks.
    \item Key findings: \todo{summarize RQ1--RQ3 results in 2--3 sentences
      each once experiments are complete.}
    \item Future work: RL training on Yedi, the 3 remaining dimensions
      (Music, Memory, Motor), multi-player/adversarial variants,
      curriculum learning across the difficulty ladder.
  \end{enumerate}
}


% ======================================================================
% Optional: Appendix
% ======================================================================
\appendix

\section{Full Configuration Table}
\plan{%
  Complete list of all 40+ benchmark configurations with active dimensions,
  sub-modes, and difficulty tier.
}

\section{Prompt Templates}
\plan{%
  The system prompt and per-turn user message templates used for LLM/VLM
  agents. Included for reproducibility.
}

\section{Per-Dimension Merge Rules}
\plan{%
  Detailed scoring tables for each dimension and sub-mode. Reference for
  readers who want to understand exactly how each multiplier tier is
  determined.
}


% ======================================================================
\bibliographystyle{plainnat}
\bibliography{references}
\plan{%
  \todo{Create references.bib with citations for:
    GamingAgent, V-MAGE, Cradle, WebArena, VisualWebArena,
    Unity ML-Agents, Gymnasium/Farama, ARC, BIG-bench,
    MORL survey, Fitts' Law, CHC cognitive theory,
    Claude/GPT-4 technical reports, MineRL, NetHack LE.}
}

\end{document}
